Sample Clustering

### 2017-07-19 Clustering.Rmd
...

## Patient info file
patient.file = grep('All_patient_info.txt', data.files, value=T)
if (length(patient.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", data))
demo.file = grep('Demographics_full.txt', data.files, value=T)

@@

patient.info = read.patient.info(patient.file, demo.file)
#head(patient.info)

patient.info = arrange(patient.info, Status, Patient, Endoscopy.Year, Pathology)

sum.patient.data = summarise.patient.info(patient.info)
sum.patient.data = as.data.frame(subset(sum.patient.data, Patient %in% names(patient.data))) ## For now
pander(sum.patient.data)

# Missing some of the samples as they aren't all sequenced yet
for (pt in names(patient.data)) {
  patient.data[[pt]]$info= patient.data[[pt]]$info[patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals)[-(1:5)],]
}

@@

## Using the segmentation values
sum.patient.data$years = with(sum.patient.data, paste(start.year, end.year, sep='-'))
sim = as.data.frame(sum.patient.data[,c('Status','years')])
rownames(sim) = sum.patient.data$Patient
for (pt in sum.patient.data$Patient) {
  print(pt)
  patient.data[[pt]]$info = arrange(patient.data[[pt]]$info,  Pathology, Endoscopy.Year)
  x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)

  gg = ggplot(melt(x1), aes(x=Var2, y=value, group=Var2, fill=Var2)) + 
    geom_boxplot(notch=T, show.legend=F) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position='none') +
    labs(title=paste(pt, patient.data[[pt]]$info$Status[1], sep=' - '), x='')

@@

@@

### Using the normalized values

So what about using the normalized segmented values for AP clustering instead?

```{r apclustN, echo=F, message=F, warning=F, fig.height=5, fig.width=5, fig.align='left'}
# Should this be performed on the normalized seg values instead?
sim3 = sum.patient.data[,c(2:4)]
rownames(sim3) = sum.patient.data[,1]
for (pt in sum.patient.data$Patient) {
  print(pt)
  x1 = patient.data[[pt]]$norm.seg.vals
  if (is.null(x1)) next
  ac = apcluster(negDistMat(x1, r=2), q=q, details=T, includeSim=T)
  #plot(ac)

@@

  for (pt in sum.patient.data$Patient[-i]) {
    x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
    ac = apcluster(negDistMat(x1, r=2), q=q)

@@

```{r xval_smp, echo=F, warning=F, message=F, fig.align='left', fig.height=4, fig.width=4}
# Leave out samples?
looS_xval = matrix(ncol=3, nrow=0, dimnames=list(c(), c('net.similarity', 'sum.similarity', 'clusters')))
for (i in 1:20) { 
  simX = sum.patient.data[,c('Status','years')]
  rownames(simX) = sum.patient.data[,'Patient']

  for (pt in sum.patient.data$Patient) {
      if (length(patient.data[[pt]]$info$Samplename) <= 2) next
      x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename[-sample(1:(length(patient.data[[pt]]$info$Samplename)), 1)])
      ac = apcluster(negDistMat(x1, r=2), q=q)

@@

simX = sum.patient.data[,c('Status','years')]
rownames(simX) = sum.patient.data[,'Patient']
# Leave out final sample on the progressors

for (pt in sum.patient.data$Patient) {
  x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
  if (subset(sum.patient.data, Patient == pt)$Status == 'P' & length(which(grepl('HGD|IMC', patient.data[[pt]]$info$Pathology))) > 0) {
    final = grep('HGD|IMC', patient.data[[pt]]$info$Pathology)

@@

subset(sum.patient.data, Patient %in% pts)

@@

pvals = as.data.frame(matrix(ncol=3,nrow=0,dimnames=list(c(), c('p.clusters','mean.clusters', 'sd.clusters'))))
status.clusters = list()
for (qv in seq(0,0.9,0.1)) {
  ## Using the segmentation values
  simPC = sum.patient.data[,c(2:4)]
  rownames(simPC) = sum.patient.data[,1]
  for (pt in sum.patient.data$Patient) {
    patient.data[[pt]]$info = arrange(patient.data[[pt]]$info, Endoscopy.Year, desc(Pathology))

>>>>

## Patient info file
patient.file = grep('All_patient_info.xlsx', data.files, value=T)
if (length(patient.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", data))
demo.file = grep('Demographics_full.xlsx', data.files, value=T)

@@

patient.info = read.patient.info(patient.file, demo.file)$info
#head(patient.info)

patient.info = arrange(patient.info, Status, Patient, Endoscopy.Year, Pathology)
patient.info = patient.info[-which(patient.info$Hospital.Research.ID == 'AH0222'),]

sum.patient.data = summarise.patient.info(patient.info)
pander(sum.patient.data)

# Missing some of the samples as they aren't all sequenced yet
#for (pt in names(patient.data)) {
#  patient.data[[pt]]$info= patient.data[[pt]]$info[patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals)[-(1:5)],]
#}

@@

## Using the segmentation values
sum.patient.data$years = with(sum.patient.data, paste(start.year, end.year, sep='-'))
sim = as.data.frame(sum.patient.data[,c('Status','years')])
rownames(sim) = sum.patient.data$Hospital.Research.ID
for (pt in sum.patient.data$Hospital.Research.ID) {
  print(pt)
  patient.data[[pt]]$info = arrange(patient.data[[pt]]$info,  Pathology, Endoscopy.Year)
  x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)

  gg = ggplot(melt(x1), aes(x=Var2, y=value, group=Var2, fill=Var2, color=Var2)) + 
    geom_boxplot(notch=T, show.legend=F) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position='none') +
    labs(title=paste(pt, patient.data[[pt]]$info$Status[1], sep=' - '), x='')

@@

```{r echo=F, message=F, warning=F, fig.width=5, fig.height=5}
g2 + labs(x='Status', y='Complexity (# clusters) per patient', title='AP Clustering', subtitle=paste("p-value", signif(ttpc, 3))) + theme(text=element_text(size=14, face='bold'), legend.position = 'non')
```

@@

```{r apclustN, echo=F, message=F, warning=F, fig.height=5, fig.width=5, fig.align='left', eval=F}
### Using the normalized values

#So what about using the normalized segmented values for AP clustering instead?

# Should this be performed on the normalized seg values instead?
sim3 = sum.patient.data[,c(2:4)]
rownames(sim3) = sum.patient.data[,1]
for (pt in sum.patient.data$Hospital.Research.ID) {
  print(pt)
  x1 = patient.data[[pt]]$norm.seg.vals
  if (is.null(x1)) next
  ac = apcluster(negDistMat(x1, r=2), q=q, details=T, includeSim=T)
  #plot(ac)

@@

  for (pt in sum.patient.data$Hospital.Research.ID[-i]) {
    x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
    ac = apcluster(negDistMat(x1, r=2), q=q)

@@

```{r xval_smp, echo=F, warning=F, message=F, fig.align='left', fig.height=4, fig.width=4, eval=F}
# Leave out samples?
looS_xval = matrix(ncol=3, nrow=0, dimnames=list(c(), c('net.similarity', 'sum.similarity', 'clusters')))
for (i in 1:20) { 
  simX = sum.patient.data[,c('Status','years')]
  rownames(simX) = sum.patient.data[,'Hospital.Research.ID']

  for (pt in sum.patient.data$Hospital.Research.ID) {
      if (length(patient.data[[pt]]$info$Samplename) <= 2) next
      x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename[-sample(1:(length(patient.data[[pt]]$info$Samplename)), 1)])
      ac = apcluster(negDistMat(x1, r=2), q=q)

@@

simX = sum.patient.data[,c('Status','years')]
rownames(simX) = sum.patient.data[,'Hospital.Research.ID']
# Leave out final sample on the progressors

for (pt in sum.patient.data$Hospital.Research.ID) {
  x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
  if (subset(sum.patient.data, Hospital.Research.ID == pt)$Status == 'P' & length(which(grepl('HGD|IMC', patient.data[[pt]]$info$Pathology))) > 0) {
    final = grep('HGD|IMC', patient.data[[pt]]$info$Pathology)

@@

subset(sum.patient.data, Hospital.Research.ID %in% pts)

@@

pvals = as.data.frame(matrix(ncol=3,nrow=0,dimnames=list(c(), c('p.clusters','mean.clusters', 'sd.clusters'))))
status.clusters = list()
for (qv in seq(0,0.9,0.1)) {
  ## Using the segmentation values
  simPC = sum.patient.data[,c(2:4)]
  rownames(simPC) = sum.patient.data[,1]
  for (pt in sum.patient.data$Hospital.Research.ID) {
    patient.data[[pt]]$info = arrange(patient.data[[pt]]$info, Endoscopy.Year, desc(Pathology))
